Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction

### 2025-01-18 src/article.tex
update related work

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order \cite{Tagliavini2007}
\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}.
Some items in the collection are not in playing condition or are too
fragile to continue to be available for the public to play. One of those
fragile instruments in the Tagliavini collection is a 1547 Harpsichord
by Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
for an interactive exhibition commisioned by San Colombano.

@@

\section{Background}\label{background}

Prvious projects looking at how to approach the problems of increasing
engagement and retaining the interction with historical musical
instruments in a museum scenario.

The paper by McAlpine \cite{McAlpine2014} also has very similar
motivations to the project presented here. Both were started from a
musical instrument museum, housing a collection of historic instruments
gifted by a collector (Benton Fletcher / Taglivaini) who wished for
collection to remain playable, which is restricted growing fragility of
some items in the respective collections. Bothe projects center around
the creation of a custioom MIDI controller for interfacing with sound
synthesis engine. The keyboard consisted of two modified commercial MIDI
keyboard controllers in a two manual harpsichord arrangement. Weherease
the McAlpine keyboard used weighted keys, this interface extends the
idea by using the mechanism and strings under tension to provide the
haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
but before taking such a leap we look at what affect a more literal
haptic keyboard may provide to the performer experience.

The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
looked at reconstructing a full tromba marina to be used as a digital
interface. The one of the limitations flagged by McAlpine was the
oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
Moderna ``attempted to avoid this problem by constructing a fully
functioning replica of the tromba marina'' \cite{Baldwin2016}. This
project applies the same approach to the harpsichord.m where the
projects deviate is a questions of scale.

Where the tromba moderna considers a single data stream this project has
had to address 98 and the problems that come with increasing scale.
Material costs mitigated by creating a scaled version of the sound
board, provinding only the length of string necessary under comparable
tension and thus resistance to the jack quill.

Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
sourcing of hardware and software for the project.

\subsection{Why\ldots{}}\label{why}

\subsubsection{\ldots a replica?}\label{a-replica}

\subsubsection{\ldots an interface?}\label{an-interface}

\begin{itemize}
\item
  maintain problems
\item
  probe for research

  \begin{itemize}
  \item
    on haptics how it affects performance
  \end{itemize}
\end{itemize}

\subsubsection{the problem}\label{the-problem}

@@

\subsubsection{the solution}\label{the-solution}

@@

\subsubsection{Criteria}\label{criteria}

A keyboard interface that provides a comparable physical response to a
harpsichord. Notes should be assigned per jack and not per key.

@@

\subsubsection{Design}\label{design}

@@

@@

\subsubsection{Fig: Construction Design
Diagram}\label{fig-construction-design-diagram}

@@

\subsubsection{Fig: Construction Design
Diagram}\label{fig-construction-design-diagram-1}

@@

Future novel designs for an interface that would utilise velocity and
aftertouch messages in software. This would be applicable for numericla
simulations of strings. A bespoke piec e of sowftare would also allow
for direct mapping of jack rows to registers. There is enough data to
implemenet functionality for velocity and aftertouch MIDI messages and
simply needs to be activated in the firmware.

Evaluation

>>>>

The project look at restoring tactile and haptic response of historical
harpsichords for an exhibition at the musical instrument Museo di San
Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
in Bologna. San Colombano is host to the Tagliavini Collection, a
collection of approximately 70 historical keyboard instruments in
working order \cite{Tagliavini2007} gifted to the museum by\\
from Luigi Ferdinando Tagliavini in 2010
\cite{SanColombano2010,Carlino2010}.

Some items among the Tagliavini collection are no longer in a playing
condition or are too fragile to continue to be available by museum
visitors. One of those fragile instruments is a 1547 Harpsichord in the
italian style by Alessandro Trasuntino
\cite{Wraight2024}\footnote{examples of similar instruments by
Trasuntino available online in the [Royal College of Music Collection
(1531)](https://museumcollections.rcm.ac.uk/collection/Details/collect/58) and
the [Musée de la musique collection
(1538)](https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView)
}, which was used as the basis for an interactive exhibition commisioned
by San Colombano.

@@

\section{Related Work}\label{related-work}

Methodologies for increasing visitor engagmemnet or

Museum staff want visitors to be engaged but limits in staffing and
funding limit how vistors can interact with collections
\cite{Templeton2018,
McAlpine2014}. For musical instrument museums there is an additional
difficulty in the balance of keeping historical instruments playable
\cite{McAlpine2013}. The fragility and decay of musical instruments
means that there is an inveitable point where instruments will no longer
be in a playable condition \cite{McAlpine2014, Fritz2017}

McAlpine outlines a similar situation to the Tagliavini collection in
his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
was a stipulation by Benton Fletcher that the instruments remain
available to play when they were gifted to National Trust Fenton House.
A custom MIDI interface arranged in a two manual harpischord style which
triggered samples of the instruments in the collection using an
appropraite recording strategy for each. One problem highlighted in user
tests by McAlpine is that the weighted keys did no provide ``an
experiential sense of interacting with a historical keyboard''
\cite{McAlpine2014}. The MIDI interface created from commercially
available weighted keys. McAlpine posits a haptic keyboard augemented
with ``actuators to provide positionally-sensitive real-time force
feedback at point-of-contact'' in a similar approach as the piano
mechanism by Gillespie \cite{Gillespie1996}.

Previous NIME project ``Tromba Moderna'' \cite{Baldwin2016} looked to
avoid the complex engineering problem by simply recreating the tromba
marina and augmenting it. In the case of the Tromba Modern, a piezo
transducer was connected to a sound synthesis engine connected to a
driver inside the instrument to simulate the vibration that would be
expected of an historical tromba marina.

The original design was to take a similar approach to the project by
Timmermans \cite{Timmermans2020}, which was an extension of the system
by Gillespie \cite{Gillespie1996}. The process laid out in Timmermans
project was to begin with a single key model of the mechanism. We began
with a 3-key Model Harpsichord Mechanism by Graziano Bandini
\textbackslash figure\{\}. Before we continued with augementing a model
with actuators it was decided that it we should first validate whether a
recreation of the harpsichord mechanism would suffice in allowing
visitors to suspend disbelief.

Benefits of putting effort into the maintenance of one interface with a
transferable sensor system

An additonal factor not faced in the Tromba Moderna project is the
problem of scale in managing 98 rather than a single data stream of
data.

Material costs mitigated by creating a scaled version of the sound
board, providing only the length of string necessary under comparable
tension and thus resistance to the jack quill.

Finally, a further extension to the projects above was a commmitement to
open sourcing all aspects of the project, hardware, software, and data.

\section{Design}\label{design}

\subsection{Design Problem}\label{design-problem}

@@

@@

The design criteria was to create a keyboard interface that provides a
comparable physical response to a harpsichord.

@@

\subsubsection{Keyboard Construction}\label{keyboard-construction}

@@

\subsection{Prototyping}\label{prototyping}

@@

@@

@@

Future novel designs for an interface that would utilise velocity and
aftertouch messages in software. This would be applicable for numericla
simulations of strings. A bespoke piec e of sowftare would also allow
for direct mapping of jack rows to registers. There is enough data to
implemenet functionality for velocity and aftertouch MIDI messages and
simply needs to be activated in the firmware. (Hyperinstruments
\cite{nime2024_20})

seat vibrations \cite{MusicalHaptics2018_07}

\begin{itemize}
\item
  ``there is a general connection between vibrations and the perceived
  quality of music reproduction''
\item
  ``influence of vibrations on loudness perception at low frequencies''
  Evaluation
\end{itemize}
